Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support).

### 2017-02-08 csf-2017/paper.tex
Reordered figure
added figure from SCM presentation
update to LR figure

@@

\subsection{Worlds}
\todo[inline]{Find a good place to describe monotonicity of
  interpretation function.}

A world is a finite map from region names, modeled as natural numbers, to regions.
We have three types of regions: \emph{permanent}, \emph{temporary}, and \emph{revoked}.
%$\perma$, $\temp$, and $\revoked$. The
%$\perma$ and $\temp$ 
Each permanent and temporary region contains a state transition system, with
public and private transitions, which describes the protocol governing
the evolution of the memory described by the region, similarly to what
has been used in logical relations for high-level
languages~\cite{Ahmed:popl09,Dreyer:jfp12,Devriese:2016ObjCap}. 
The protocols imposed on memory by permanent regions stay in place
indefinitely. This means that any capability, local or global, can
depend on these protocols. The protocols imposed on memory by temporary
regions can be revoked at certain points during execution,
corresponding to the points when local capabilities can be
revoked. This means that local capabilities safely can depend on the
protocols imposed on memory by temporary regions, wheareas global
capabilities cannot, since a global capability may stay around
throughout the execution of a program, but a temporary region may be
revoked.

@@

\subsection{Logical relation}

@@

\begin{figure}[htbp]
  \centering
\begin{align*}
  & \observations \; : \;  \Worlds \nefun \UPred{\Regs \times \HeapSegments} \\
  & \observations (W) \defeq \\
  & \;\left\{ \npair{(\reg,\ms)} \; \middle| \;
    \begin{aligned}
      & \forall \ms_f, \heap', i \leq n \ldotp \\
      & (\reg,\ms \uplus \ms_f) \step[i] (\halted,\heap')  \\
      & \;\; \Rightarrow
      \begin{aligned}[t]
        & \exists W' \futurestr W, \ms_r, \ms' \ldotp\\
        & \;\; \heap' = \ms' \uplus \ms_r \uplus \ms_f \text{\footnotesize{ and }} \\ 
        & \;\; \heapSat[\hs']{n-i}{W'} 
      \end{aligned}
    \end{aligned}\right\}\\
  & \stdrr \; : \; \Worlds \monwknefun \UPred{\Regs} \\
  & \stdrr(W) \defeq \left\{ \npair{\reg} \; \middle| \;
                  \begin{aligned}
                      & \forall r \in \RegName \setminus \{\pcreg\} \ldotp \\
                    & \;\;  \npair{\reg(r)} \in \stdvr(W) 
                  \end{aligned} \right\} \\
  & \\
  & \stder \; : \; \Worlds \nefun \UPred{\Words} \\
  & \stder(W) \defeq \left\{ \npair{\pc} \; \middle| \; 
                  \begin{aligned}
                    & \forall n' \leq n, \npair[n']{\reg} \in \stdrr(W), \heapSat[\hs]{n'}{W} \ldotp \\
                    & \;\;\npair[n']{(\reg\update{\pcreg}{\pc},\hs)} \in \observations(W) 
                  \end{aligned} \right\} \\
  & \\
  &\stdvr \; : \; \Worlds \monwknefun \UPred{\Words} \\
  &\stdvr(W)\defeq  \\
    &\;\;\begin{aligned}[t]
      & \{ \npair{i} \mid i \in \ints \} 
      \union \\
      & \{ \npair{\stdcap[(\noperm,\gl)] }  \} 
      \union \\
      & \{ \npair{\stdcap[(\readonly,\gl)] } \mid \\
      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)\} 
      \union \\
      & \{ \npair{\stdcap[(\readwrite,\gl)] } \mid \\
      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
      & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \}
      \union \\
      & \{ \npair{\stdcap[(\readwritel,\gl)] } \mid \\
      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
      & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \}
      \union \\
      & \{ \npair{\stdcap[(\exec,\gl)]} \mid \\
      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
      & \quad \npair{(\{\exec\},\start,\addrend)} \in \execCond{}(\gl)(W) \} 
      \union \\
      & \{ \npair{\stdcap[(\entry,\gl)]} \mid \\
      & \quad \npair{(\start,\addrend,\addr)} \in \entryCond{}(\gl)(W)\} 
      \union \\
      & \{ \npair{\stdcap[(\rwx,\gl)]} \mid \\
      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
      & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \text{\footnotesize{ and }}\\
      & \quad \npair{(\{\rwx,\exec\},\start,\addrend)} \in \execCond{}(\gl)(W) 
      \union \\
      & \{ \npair{\stdcap[(\rwlx,\gl)]} \mid \\
      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
      & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \text{\footnotesize{ and }}\\
      & \quad \npair{(\{\rwlx,\rwx,\exec\},\start,\addrend)} \in \execCond{}(\gl)(W) \}
    \end{aligned}
\end{align*}
  \caption{Logical Relation}
  \label{fig:logrel}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \begin{align*}
    &\iota = (v,s,\phi_\pub,\phi,H) \text{\footnotesize{ is address-stratified iff}}\\
    &\qquad\begin{multlined}
      \forall s', \hat{W}, n, \ms, \ms' \ldotp \\
      \npair{\ms},\npair{\ms'} \in H~ s'~ \hat{W} \Rightarrow \\
      \dom(\ms) = \dom(\ms') \wedge \\
      \forall \addr \in
      \dom(\ms)\ldotp \npair{\ms\update{\addr}{\ms'(\addr)}} \in H~ s'~ \hat{W}
    \end{multlined}
  \end{align*}
  \begin{align*}
  & \writeCond{}(\iota,\gl)(W) =  \\
  & \left\{ \npair{(\start,\addrend)} \; \middle| \;
    \begin{aligned}
      & \exists r \in \var{localityReg}(g,W) \ldotp \\
      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
      & \;\;\;\; W(r)\nsupsim[n-1] \iota_{\start',\addrend'} \text{\footnotesize{ and}}\\
      & \;\;\;\; W(r) \text{\footnotesize{ is address-stratified}} 
    \end{aligned} \right\}\\
  & \readCond{}(\gl)(W) =  \\
  & \left\{ \npair{(\start,\addrend)} \; \middle| \;
    \begin{aligned}
      & \exists r \in \var{localityReg}(g,W) \ldotp \\
      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
      & \;\;\;\; W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl 
    \end{aligned} \right\}\\
  & \execCond{}(\gl)(W) = \\
  & \left\{ \npair{(\var{perms},\start,\addrend)} \; \middle| \;
    \begin{aligned}
      & \forall n' < n, W' \future W \ldotp \\
      & \;\; \forall a \in [\start,\addrend], \perm \in \var{perms} \ldotp \\
      & \;\;\;\; \npair[n']{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W')\\
    \end{aligned} \right\} \\
  & \entryCond{}(\gl)(W) = \\
  & \left\{ \npair{(\start,\addrend,\addr)} \; \middle| \;
    \begin{aligned}
 &  \forall n' < n \ldotp \forall W' \future W \ldotp\\
      & \quad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')
    \end{aligned} \right\} \\
  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
  & \quad \text{and } \gl = \glob \Rightarrow \future = \futurestr
\end{align*}

\caption{Permission-based conditions}
\label{fig:perm-conds}
\end{figure}
Now define the value relation as follows:

@@

\bibliographystyle{plainnat}
\bibliography{references}
\end{document}
\grid

>>>>

\usepackage{afterpage}

@@

\subsection{Worlds}
\todo[inline]{Find a good place to describe monotonicity of
  interpretation function.}
\begin{figure}[htbp]
  \centering
  \includegraphics{w11}
  \caption{Relation between local and global capabilities and temporary and permanent regions.}
  \label{fig:cap-world}
\end{figure}
A world is a finite map from region names, modeled as natural numbers, to regions.
We have three types of regions: \emph{permanent}, \emph{temporary}, and \emph{revoked}.
%$\perma$, $\temp$, and $\revoked$. The
%$\perma$ and $\temp$ 
Each permanent and temporary region contains a state transition system, with
public and private transitions, which describes the protocol governing
the evolution of the memory described by the region, similarly to what
has been used in logical relations for high-level
languages~\cite{Ahmed:popl09,Dreyer:jfp12,Devriese:2016ObjCap}. 
The protocols imposed on memory by permanent regions stay in place
indefinitely. This means that any capability, local or global, can
depend on these protocols. The protocols imposed on memory by temporary
regions can be revoked at certain points during execution,
corresponding to the points when local capabilities can be
revoked. This means that local capabilities safely can depend on the
protocols imposed on memory by temporary regions, wheareas global
capabilities cannot, since a global capability may stay around
throughout the execution of a program, but a temporary region may be
revoked.

@@

\subsection{Logical relation}
\afterpage{
\begin{figure}[htbp]
  \centering
\begin{align*}
  & \observations \; : \;  \Worlds \nefun \UPred{\Regs \times \HeapSegments} \\
  & \observations (W) \defeq \\
  & \;\left\{ \npair{(\reg,\ms)} \; \middle| \;
    \begin{aligned}
      & \forall \ms_f, \heap', i \leq n \ldotp \\
      & (\reg,\ms \uplus \ms_f) \step[i] (\halted,\heap')  \\
      & \;\; \Rightarrow
      \begin{aligned}[t]
        & \exists W' \futurestr W, \ms_r, \ms' \ldotp\\
        & \;\; \heap' = \ms' \uplus \ms_r \uplus \ms_f \text{\footnotesize{ and }} \\ 
        & \;\; \heapSat[\hs']{n-i}{W'} 
      \end{aligned}
    \end{aligned}\right\}\\
  & \stdrr \; : \; \Worlds \monwknefun \UPred{\Regs} \\
  & \stdrr(W) \defeq \left\{ \npair{\reg} \; \middle| \;
                  \begin{aligned}
                      & \forall r \in \RegName \setminus \{\pcreg\} \ldotp \\
                    & \;\;  \npair{\reg(r)} \in \stdvr(W) 
                  \end{aligned} \right\} \\
  & \\
  & \stder \; : \; \Worlds \nefun \UPred{\Words} \\
  & \stder(W) \defeq \left\{ \npair{\pc} \; \middle| \; 
                  \begin{aligned}
                    & \forall n' \leq n, \npair[n']{\reg} \in \stdrr(W), \heapSat[\hs]{n'}{W} \ldotp \\
                    & \;\;\npair[n']{(\reg\update{\pcreg}{\pc},\hs)} \in \observations(W) 
                  \end{aligned} \right\} \\
  & \\
  &\stdvr \; : \; \Worlds \monwknefun \UPred{\Words} \\
  &\stdvr(W)\defeq  \\
    &\;\;\begin{aligned}[t]
      & \{ \npair{i} \mid i \in \ints \} 
      \union \\
      & \{ \npair{\stdcap[(\noperm,\gl)] }  \} 
      \union \\
      & \{ \npair{\stdcap[(\readonly,\gl)] } \mid \\
      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)\} 
      \union \\
      & \{ \npair{\stdcap[(\readwrite,\gl)] } \mid \\
      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
      & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \}
      \union \\
      & \{ \npair{\stdcap[(\readwritel,\gl)] } \mid \\
      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
      & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \}
      \union \\
      & \{ \npair{\stdcap[(\exec,\gl)]} \mid \\
      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
      & \quad \npair{(\{\exec\},\start,\addrend)} \in \execCond{}(\gl)(W) \} 
      \union \\
      & \{ \npair{\stdcap[(\entry,\gl)]} \mid \\
      & \quad \npair{(\start,\addrend,\addr)} \in \entryCond{}(\gl)(W)\} 
      \union \\
      & \{ \npair{\stdcap[(\rwx,\gl)]} \mid \\
      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
      & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \text{\footnotesize{ and }}\\
      & \quad \npair{(\{\rwx,\exec\},\start,\addrend)} \in \execCond{}(\gl)(W) 
      \union \\
      & \{ \npair{\stdcap[(\rwlx,\gl)]} \mid \\
      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }} \\
      & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \text{\footnotesize{ and }}\\
      & \quad \npair{(\{\rwlx,\rwx,\exec\},\start,\addrend)} \in \execCond{}(\gl)(W) \}
    \end{aligned}
\end{align*}
  \caption{Logical Relation}
  \label{fig:logrel}
\end{figure}

\begin{figure}[htbp]
  \centering
  \begin{align*}
  & \writeCond{}(\iota,\gl)(W) =  \\
  & \left\{ \npair{(\start,\addrend)} \; \middle| \;
    \begin{aligned}
      & \exists r \in \var{localityReg}(g,W) \ldotp \\
      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
      & \;\;\;\; W(r)\nsupsim[n-1] \iota_{\start',\addrend'} \text{\footnotesize{ and}}\\
      & \;\;\;\; W(r) \text{\footnotesize{ is address-stratified}} 
    \end{aligned} \right\}\\
  & \readCond{}(\gl)(W) =  \\
  & \left\{ \npair{(\start,\addrend)} \; \middle| \;
    \begin{aligned}
      & \exists r \in \var{localityReg}(g,W) \ldotp \\
      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
      & \;\;\;\; W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl 
    \end{aligned} \right\}\\
  & \execCond{}(\gl)(W) = \\
  & \left\{ \npair{(\var{perms},\start,\addrend)} \; \middle| \;
    \begin{aligned}
      & \forall n' < n, W' \future W \ldotp \\
      & \;\; \forall a \in [\start,\addrend], \perm \in \var{perms} \ldotp \\
      & \;\;\;\; \npair[n']{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W')\\
    \end{aligned} \right\} \\
  & \entryCond{}(\gl)(W) = \\
  & \left\{ \npair{(\start,\addrend,\addr)} \; \middle| \;
    \begin{aligned}
 &  \forall n' < n \ldotp \forall W' \future W \ldotp\\
      & \quad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')
    \end{aligned} \right\} \\
  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
  & \quad \text{and } \gl = \glob \Rightarrow \future = \futurestr
\end{align*}
  \begin{align*}
    &\iota = (v,s,\phi_\pub,\phi,H) \text{\footnotesize{ is address-stratified iff}}\\
    &\qquad\begin{multlined}
      \forall s', \hat{W}, n, \ms, \ms' \ldotp \\
      \npair{\ms},\npair{\ms'} \in H~ s'~ \hat{W} \Rightarrow \\
      \dom(\ms) = \dom(\ms') \wedge \\
      \forall \addr \in
      \dom(\ms)\ldotp \npair{\ms\update{\addr}{\ms'(\addr)}} \in H~ s'~ \hat{W}
    \end{multlined}
  \end{align*}

\caption{Permission-based conditions}
\label{fig:perm-conds}
\end{figure}
\clearpage
}

@@

@@

Now define the value relation as follows:

@@

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,references}
\end{document}
\grid
